---
title: "R scripts for 'Discovering Active Subpaces for High Dimensional Computer Models'"
output: github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


This repository contains several R scripts to reproduce the examples found in the (submitted for review) manuscript *Active Subspaces for High Dimensional Computer Models*.

To run these scripts, you will need to install the `concordance` and `GBASS` packages (other packages from CRAN may be neeed, see `R/setup.R` for details) from public github repos using the commands.

```{r, eval = F}
# install.packages("devtools")
devtools::install_github("knrumsey/concordance")
devtools::install_github("knrumsey/GBASS")
```

A link to the manuscript will be included when available. A Quarto document (and the rendered pdf) titled `Examples_Walkthrough` includes a demonstration of the examples from sections 4.1 and 4.2 in the manuscript. Additional scripts are included to reproduce the results included in Sections 4.1-4.2 as well as some of the results in the Supplemental materials. The scripts included in this repository include

   - `A_simple_polynomial.R` (Section 4.1)
   - `B_linear_constraint.R` (Section 4.1.1)
   - `C_low_dim_structure.R` (Section 4.2)
   - `D_polynomial_outliers.R` (Supplemental materials)
   - `E_polynomial_quantile.R` (Supplemental materials)
   


# References
